---
output:
  html_document:
    css: styles.css
---

<br>
<br>
<br>

# How Sustainable Is Your State? 
<br>

#### Investigating The United States' Global Environmental Impact. {#title2}
<br>
<br>

by Alison McGuire, Alyssa Holzer and MacKenzie Olson
<br>
<br>

###### Graph 1 
###### The States' Overall Impact Scores and Ranks 
```{r, fig.align='center', out.extra='angle=90',eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
source("script.R")
overall.map 
```

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

## Table of Contents
[Introduction](#introduction)

[Literature Review](#literature-review)

[Method Description](#method-description)

[Data Analysis](#data-analysis)

[Discussion](#discussion)

[Reference List](#reference-list)

[Appendix](#appendix)

<hr>

## Introduction

Environmental sustainability is the practice of using renewable and environmentally friendly practices in order to protect the earth's resources and provide for future generations. In this report we seek to understand the environmental impact of every state by analyzing data on landfills, water withdrawals, and power plant emissions for each state. As organic materials in landfills decompose they emit ozone depleting substances, also known as landfill gas (LFG), which have a negative effect on the environment and contribute to global warming. However, some landfills have LFG systems in place that collect emissions and lessen the amount of Greenhouse Gases emitted. Water withdrawals refer to the amount of water withdrawn from a natural groundwater or surface water source. There are many activities such as agriculture, power generation, and manufacturing that require water and thus contribute to water withdrawals across the country. Water withdrawal has many associated problems including drought. That being said, more efficient transfer of water such as storm water capture and water recycling can help reduce the impact of water withdrawals. Power plants use a variety of fuels to produce energy, including fossil fuels such as coal, oil, and natural gas. When these fuels are burned they produce Greenhouse Gases, which contribute to global warming. Other renewable sources of energy such as solar, wind, hydro, and geothermal do not emit Greenhouse Gases, and are therefore much better for the environment. Different states may have varying laws and regulations regarding these three areas, and therefore have varying levels of environmental impact. Using data on landfills, water withdrawals and power plant emissions, we have created scores for each of the 50 states in order to rank each's  environmental impact.

<hr>

## Literature Review

### Landfills 

In _The municipal solid waste landfill as a source of ozone-depleting substances in the United States and United Kingdom_ by E.L. Hodson, D. Martin, and R. G. Prinn, said researchers looked to better understand the relationship between lingering ozone depleting substance (ODS) emissions and landfills who emit them. It is known fact that ODS influence climate change as they weaken the Earth’s defense against the sun: the ozone. When waste is not disposed of properly in landfills, it has the potential to release ODS.

Luckily, the accumulation of ODS has depleted to the point where tropospheric concentrations are stable or decreasing (Hodson et. all, 2010, pg. 1). However, accurate predictions of future ODS emissions is necessary so that appropriate strategies are developed to minimize ozone loss. While there are several different types of ODS, for the sake of simplicity Hodson et. all decided to study the following four: CFC- 11 (trichlorofluoromethane), CFC-12 (dichlorodifluoromethane), CFC-113 (1,1,2-trichloro-1,2,2-trifluoroethane), and CH3CCl3 (1,1,1-trichloroethane) (Hodson et. all, 2010, pg. 1). These four were selected because of their high ozone depleting potential (ODP) and the fact they were historically released in large quantities into the atmosphere. Furthermore, they have long lifetimes until they dissolve which further increases their impact. Prior to this study, CFC-11, CFC-12, CFC-1143, and Ch3CCI3 have all been detected at above normal levels in landfills (Hodson et. all, 2010, pg. 2). Despite this compelling evidence that these ODS need to be monitored, there are not any existing national inventory methods to quantify national ODS emissions from landfills (Hodson et. all, 2010, pg. 2). 

In terms of methodology, Hodson et. all studied seven US and nine UK municipal solid waste (MSW) landfills with active gas management (2010, pg. 2). It is important to note that these landfills accept more than 50% of their waste from domestic and commercial sources (Hodson et. all, 2010, pg. 2). These landfills additionally varied in size- ranging from small, medium and large- to get the most representative sample. Most fields received one-time field sampling; however, one US MSW landfills was sampled every month for over a year (Hodson et. all, 2010, pg. 2).  The US MSW landfills had triplicate canister samples where they UK MSW landfills only had one canister samples. 

After performing various chemistry based calculations to get the different percentages of gases that made up the various samples, Hodson et. all developed confidence intervals to best estimate the percentage these CFCs and CH3CCI3 make-up of the landfill emissions.  The results were the following: CFC-11 (0.037 - 0.074 Gg/y), CFC-12 (0.089 - 0.18 Gg/y), CFC-113 (0.0058 - 0.012 Gg/y), CH3CCl3 (0.012 - 0.024 Gg/y) (Hodson et. all, 2010, page 7). Understandably, the results without context do not mean much. Luckily, these quantities support the hypothesis that US and MK MSW landfills aren’t a significant source of these gases. In fact, these estimates indicate that US MSW landfills account for less than 1% of total US emissions (Hodson et. all, 2010, page 10). Thus, this study shows that monitoring and properly disposing of trash can keep ODS levels low and as a result make a significant difference. 

### Water withdrawals

In _Factors Governing Change in Water Withdrawals for U.S. Industrial Sectors from 1997 to 2002_, Hui Wang, Mitchell J. Small, and David A. Dzomback looked to understand why water withdrawal levels have been maintained despite a rapidly growing population. Wang et. all defined water shortage as the total amount of water withdrawn from its source (2014, pg. 1). Water withdrawal can be alleviated through strategies like “water recycling and reuse, storm water capture, water transfer from more water-rich areas, and desalination of seawater” (Wang et. all, 2014, paragraph 1).  It was found that in 2002, US sectors like agriculture, power generation, and food manufacturing were responsible for 50% of the water withdrawal (Wang et. all, 2014, paragraph 2). As the population grows, it relies more heavily on these basic sectors and can impact water quantity, quality, and availability. Luckily, despite population growth rates of water withdrawal have been increasing at a much slower place. Inspired by this phenomena, Wang et. all decided to study 5 factors: “changes in population, GDP per capita, water use intensity, production structure, and consumption patterns” (Wang et. all, 2014, paragraph 4). 

In terms of data collection, Wang et. all used the U.S. Bureau of Economic Analysis (BEA) benchmark economic input-output (EIO) table from 1997 and 2002 to gain information about water withdrawals about summary sectors for both years (Wang et. all, 2014, paragraph 5). Wang et. all additionally used USGS data to get information about Thermoelectric power generation, Mining, Industrial, Irrigation, Residential, Public supply, Livestock, and Aquaculture (Wang et. all, 2014, paragraph 6). 

Once the data was collected, Wang et. all created an Economic Input-Output Life Cycle Assessment (EIO-LCA). This uses independent variables that are subsections of the 5 main factors listed above to estimate the overall water withdrawal. From there, a structural decomposition analysis (SDA) was done to see how the variables independent variables influence the dependent variable as well as their relationship with each other. Upon extensive regression-based modeling, Wang et. all found that water use intensity was able to most accurately predict water withdrawal rates (Wang et. all, 2014, paragraph 34). Thus, it was important to keep water withdrawal rates in mind when examining state to state.
	
### Power Plants Emissions

_North American Power Plant Air Emissions_ describes the Commission for Environmental Cooperation’s (CEC) 2011 report on criteria air pollutant and Greenhouse Gas emissions from fossil fuel-fired power plants in North America for the year 2005. This was the most recent year for which data was available from United States, Canada, and Mexico when the report was written. For the purpose of our research, we focused on their analysis of data collected on power plants in the United States. Criteria air pollutants (including sulfur dioxide, and nitrogen oxides) and Greenhouse Gases are atmospheric pollutants emitted by power plants. It is also possible for power plants to release trace metals, such as mercury, with certain fuels (Anonymous, 2012, page 1.1). About two-thirds of power generation in North America comes from fossil fuels including coal, residual fuel oil, and natural gas, which release air pollutants when burned (Anonymous, 2012, page 1.2).

The CEC’s methodology involved collecting public information on individual fossil fuel-fired power plants, including their installed capacity, electricity generation, technologies used, and fuels burned from public sources including each country’s national emission inventories. They used other public information to calculate estimates where data was absent. Limited availability of data narrowed the CEC’s analysis to data on SO2, NOX, and CO2 emissions (Anonymous, 2012, page 1.1). For the United States, they obtained data for CO2, NOX, SO2, methane, nitrous oxide and mercury emissions from 2,728 registered facilities. Data on PM-10 (particulate matter with a diameter of 10 mircormeters or less) and PM-2.5 (particulate matter with a diameter of 2.5 micrometers or less) emissions were only included for 1,182 of the facilities (Anonymous, 2012, page 1.2).

In their analysis, the CEC determined there was no significant change in the overall amount of emissions across North America between 2002 and 2005 (Anonymous, 2012, 1.3). In 2005, the net electricity generation in the U.S. was 49.7% from coal, 18.7% from natural gas, 5.46% from hydroelectric, 2.3% from renewable energy sources (including biomass, wind, geothermal, and solar), and 3% from other sources (Anonymous, 2012, pages 2-3). Electricity generation is the source of about one-third of total Greenhouse Gas emissions in the U.S, contributing 39.3% of CO2 emissions nation-wide in 2005 (Anonymous, 2012, page 3). The CEC also found that coal-fired power plants were the largest contributor of SO2, NOx, PM-10, PM-2.5, and Greenhouse Gas emissions in the U.S. (Anonymous, 2012, page 5).

### Summary
Now that a sufficient background has been laid down about waste disposal, water withdrawal, and power consumption, this paper will delve into its area of focus. It is clear that extensive research has been in all of these areas individually; however, with a problem as large as climate change, an interdisciplinary approach is required to get a holistic overview of the problem without compromising the quality of information. As a result of this belief, this paper takes on a pragmatic worldview, constructed by a statistician, data-visualization expert, and a data analyst. All of these papers offer great depth, however, this research article will focus on how we assigned each of the 50 United States a score based on how they are navigating their relationship between waste disposal, power consumption, and water withdrawals. 


<hr>

## Method Description


### Introduction
Our dataset consists of three major categories that contribute to environmental hazards: landfills, power plants, and water withdrawal. We have aggregated and cleaned a few different datasets to fit our needs and the data now lives inside one CSV document. There is quantitative, state by state data on landfills, waste, power consumption, emissions, water withdrawals, and water waste. 


### Landfills

##### Source
This data is from the Environmental Protection Agency’s (EPA) Landfill Methane Outreach Program (LMOP) and is current as of June 2017. The LMOP is an program that aims to reduce methane emissions from landfills. One of the ways they accomplish this goal is through gathering data on landfills and potential energy projects. 

##### Description
This data has two major parts, information about the itself landfill and information about methane saving projects at the landfill ("Access Landfill", 2017). The landfill data is a collection of 2,400 municipal solid waste landfills across the United States. This data has been collected largely in part due to the partnerships the EPA has with landfills. It includes landfills that both do and do not report to the EPA’s Greenhouse Gas Reporting Program. The landfill specific data includes information like the state it is located in, the landfill’s name, the landfill operator’s name (or organization), the county it is located in, the latitude and longitude of the plant and the amount of trash it contains (in tons). It also contains information about landfill operations including the year it opened, the year it has closed applicable, if they have methane protection projects going on, the year those methane protection projects started, and much more. The appendix has all the columns from this dataset. 

### Power Plant Emissions

##### Source
This data has also been collected by the EPA and is current as of 2014. It is a part of the EPA’s Emissions & Generated Resource Integrated Database (“eGRID”, 2017). The EPA describes eGRID as “a comprehensive source of data on the environmental characteristics of almost all electric power generated in the United States” (“eGRID”, 2017). They collect data on air emissions for nitrogen oxides, sulfur dioxide, carbon dioxide, methane, and nitrous oxide; emissions rates; net generation; resource mix; and many other attributes.

There are three major sources that the eGRID gets its data from:
<br>
1.	EPA/CAMD: this includes data reported to EPA by electric generating units to comply with the regulations in 40 CFR Part 75. Data include annual emissions of CO2, NOx, and SO2; ozone season emissions of NOx; and annual and ozone season generation and heat input. The data is available at https://www.epa.gov/airmarkets/.
2.	EIA-860: this includes data reported to EIA on electric generators. Data include nameplate capacity, prime mover, primary fuel type, and indication of whether the generator is a combined-heat-and-power unit (EIA, 2016a).
3.	EIA-923: this includes data reported to EIA on fuel consumption and generation. Data include monthly generation and heat input at the unit or generator level for a subset of units and generators, and at the prime mover level for all plants. eGRID2014 uses unit- or generator-level data where available, and prime mover-level data for all other units and generators (EIA, 2016b).


##### Description
This dataset comes with hundreds of columns; we have narrowed down those to just a few that are important to our mission. It describes the different amounts of emissions from various Greenhouse Gases as well as the amount and type of energy it produces. It also shows summary statistics on the amount of renewable versus nonrenewable energy and combustible versus noncombustible energy that it generates.


### Water Withdrawals

##### Source
The data was gathered from the United States Geological Survey (“Misson”). This agency collects, monitors and assesses data on different aspects of the United States’ geology for both policy makers and public interest. Their data contains information about many different aspects including land, air, water and everything in between. This data is current as of 2010. 

##### Description
This data describes how much water is being used by the population of the United States. The data has been aggregated to reflect a state by state total for each column. Furthermore, it has been separated into many different parts, including where the water is coming from (surface water or groundwater), where it’s going to (agriculture, mining, industrial, livestock, aquaculture), who is using it (publicly supplied, domestically supplied), and the type of water (fresh or saline). There is also information on the populations for each state and the populations that use this water. 

### Analysis
We performed multiple regression analysis to see if there is a linear, predictive relationship between the rankings we assigned to the states (dependent variable) and the scores that compose said ranking (independent variables) which are built on the following 6 categories: total emissions, total water withdrawals, total waste, percentage renewable, total LFG collected, and percentage non-combustible. The scoring system will be described in detail later below. This test was done to make sure how we weighted and created the scores aligns well with the overall ranking system. There are 3 checks we used to make sure multiple regression is an appropriate model to use: 

1. There is a linear relationship between the independent and dependent variables.
2. For each value of X, the probability distribution of Y has the same standard deviation and the y values are independent.
3. The dependent variable’s distribution is roughly normally distributed.

<br>

<hr>

## Data Analysis

### Introduction

Our analysis revolved around six variables that we chose to represent the environmental impacts of the different states. The six variables were chosen out of our three major datasets (Water Withdrawals, Landfills, and Power Plants). First, we looked at the total amount of emissions in tons per capita per state. Secondly, we took the total water withdrawals in millions of gallons per day as a percentage of the population in each state. Thirdly, we looked at the amount waste in tons per person for each state. The fourth variable was the percentage of renewable energy vs the total energy produced by each state. The fifth variable was amount of emissions collected through an LFG system for each state. The last variable we took into consideration was the ratio of combustible fuel vs noncombustible fuels for each state.  

Here is a summary of how each score was calculated. 
<br>
For the Waste score the calculation is as follows:
$$ Waste \space Score = 1 - {{State \space Total \space Waste \space Amount \space (tons) * (100 - Percent \space Landfills \space With \space No \space Waste \space Data) \over State Population } \over Largest  \space Amount  \space Of  \space Waste \space Generated  \space By  \space All  \space  50  \space  States } $$
For the LFG score the calculation is as follows:
$$ LFG \space Score = {{State \space Total \space LFG \space Collected \space (mmscfd) \over State Population} \over Largest  \space Amount  \space Of  \space LFG \space Collected \space Per \space Capita \space Of  \space All  \space  50  \space  States } $$
For the Noncombustibles score the calculation is as follows:
$$ Noncombustibles \space Score = {{State \space Total \space Noncombustibles \over State \space Total \space Net \space Generation } \over Largest  \space Percent  \space Of  \space Noncombustibles \space Generated \space Of  \space All  \space  50  \space  States } $$
For the Renewables score the calculation is as follows:
$$ Renewables \space Score = {{State \space Total \space Renewables \over State \space Total \space Net \space Generation } \over Largest  \space Percent  \space Of  \space Renewables \space Generated \space Of  \space All  \space  50  \space  States } $$
For the Emissions score the calculation is as follows:
$$ Emissions \space Score = 1 - {{State \space NOx \space Emissions (tons) + State \space CO2 \space Emissions (tons) + State \space SO2 \space Emissions (tons) + \\ (State \space CH4 \space Emissions (lbs) * 0.005) + (State \space N2O \space Emissions (lbs) * 0.005) \over State \space Total \space Net \space Generation } \over Largest  \space Amount  \space Of  \space Emissions \space Generated \space Of  \space All  \space  50  \space  States } $$
For the Water score the calculation is as follows:
$$ Water \space Score = {{State \space Total \space Water \space Withdrawals \space (Mgal/d) \over State Population} \over Largest  \space Amount  \space Of  \space Water \space Withdrawals \space Per \space Capita \space Of  \space All  \space  50  \space  States } $$


<br>
We knew that these variables did not have the same amount of impact so the next step was to determine which variables should be weighed more heavily. We consulted an expert in the field, Lynn McGuire, as well as our own literature review in order to determine how we would weigh each variable. Our research concluded that they should be placed in the following order:

1. Total emissions
2. Total water withdrawals
3. Total waste
4. Percentage of renewable energy
5. Total LFG collected
5. Percentage of non-combustibles

Total emissions, water withdrawals and waste were all ranked near the top because they were the most direct impacts on the environment. Greenhouse Gas emissions have been proven to directly cause climate change on Earth, lending this to be the first in our rankings of variables. Water withdrawals also directly contribute to the earth's climate problem because it is measuring the impact humans are having on the water ecosystems. Waste comes in third because it is both influencing ecosystems by poisoning land and contributing to the Greenhouse Gas emissions. Renewable energy and LFG collection are important to the prevention of Greenhouse Gas emissions. Finally, the percentage of non-combustible gasses which have potential to be dangerous to our earth, but are not directly influencing climate change. 
 
Once we ranked the variables we decided on a weighted system to determine each state's score. We weighted the more important variables heavier than the less important (based on our rankings). These weights are by no means the single, correct solution, but rather what we found to be a fair system based off of our literature review. A future research goal could be experimenting with various distributions of this weighting system and the resulting impact on the scores. As of now, the total score comprised of:

25% : Total emissions <br>
20% : Water Withdrawals <br>
16% : Total waste <br>
14% : Percentage of renewable energy <br>
13% : Total LFG collected <br>
12% : Percentage of non-combustibles <br>

These factors were tehen summed to composed our total score for each state. The states were then assigned a rank depending on their score. A higher score correlated to a higher rank meaning the state had a low environmental impact. The rest of our analysis was completed based on this score. 

### Summary Statistics

This table includes the scores for each of our focus categories for each state, as well as the total score for each state. 
```{r eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library(knitr)
kable(all.data.output, caption = "Table 1")

```

<br>
Here is a list of some summary statistics gathered from the scores above. 


<br>
```{r eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

names(sum.stats)<-c("Variable", "Mean", "Median")
kable(sum.stats, caption = "Table 2")


```
<br>

###### Graph 2
###### Plot of Scores by State and Score Type
```{r eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
distribution.plot
```


### Multivariate Regression

We conducted a multivaraite regression test to check that our scores were assigned to the states' ranks in an appropiate and linear fashion.

#### Statistical Checks

The results of the checks, mentioned earlier in this paper, to make sure multiple regression is an appropriate test are as follows:

##### 1.	There is a linear relationship between the independent and dependent variables. 

Unfortunately, there did not seem to be a strong linear relationship in the scatterplots above. However, we did not see any different types of patterns such as logarithmic or exponential functions which is even more worrisome. As a result, we proceeded cautiously. 
<br>

###### Graph 3
###### Scores Given Emissions
```{r eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
source("linear_regression.R")
a
```
<br>

###### Graph 4
###### Scores Given Withdrawals
```{r eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

b
```
<br>

###### Graph 5
###### Scores Given Waste by Population
```{r eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

c
```
<br>

###### Graph 6
###### Scores Given LFG Collected
```{r eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

d
```
<br>

###### Graph 7
###### Scores Given Renewables
```{r eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

e
```
<br>

###### Graph 8
###### Scores Given Combustion
```{r eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

f
```
<br>

##### 2. For each value of X, the probability distribution of Y has the same standard deviation and the y values are independent. 

The easiest way to evaluate this check is to check the graphs of the residuals given the fitted value. The variability of the residuals should fairly consistent given any value of x and there shouldn’t be a clear pattern. 
<br>

###### Graph 9
###### Emission Score Residuals vs Fitted Values
```{r eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

p.emissions

```
<br>

###### Graph 10
###### Withdrawal Score Residuals vs Fitted Values
```{r eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

p.withdrawals

```
<br>

###### Graph 11
###### Waste Score Residuals vs Fitted Values
```{r eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

p.waste.pop

```
<br>

###### Graph 12
###### LFG Score Residuals vs Fitted Values
```{r eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

p.lfg.collected

```
<br>

###### Graph 13
###### Renewables Score Residuals vs Fitted Values
```{r eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

p.renews

```
<br>

###### Graph 14
###### Combustion Score Residuals vs Fitted Values
```{r eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

p.combustion
```


As displayed by the graph, it seems to be an okay fit! There isn’t a clear pattern to the residuals and they appear to vary by approximately the same amount. 

##### 3. The dependent variable’s distribution is roughly normally distributed

As one can see in Graph 15 below, there is a uniform distribution for scores. This is not a normal distribution; however, there is a lack of skew or signs of being bimodal and thus we will proceed with caution.


###### Graph 15
###### Distribution of Scores by State
```{r eval=TRUE, echo=FALSE, warning=FALSE, evaluted=FALSE, error=FALSE, message=FALSE}

h
```


### Results

The following equation was created by the linear regression function in R: 
$$ Predicted\space Rank = 85.655 - 39.127 (emissions) - 13.994 (withdrawals) - 24.256 (waste) – 12.535 (combustion) - 15.064 (lfg.collected) – 6.114 (renews)$$

By creating a linear regression function in R, R additionally runs an F-test. The null hypothesis for an f-test is that a model with only a y-intercept is just as good a fit as the model created through linear regression. After conducting the f-test, we received a sample statistic of 109.9 If we were to conduct a million f-tests with a million different randomly selected samples, the odds of getting a more extreme f-statistic is ~0.0002%. As a result, we can reject the null hypothesis. This means that we have created a model that is better than a model with only the y-intercept.

Overall, this confirms that we have made a model for assigning ranks that is fairly predictable which is encouraging. This means that if one only knew information about the 6 independent variables, they could use this equation to get a predicted rank that would be extremely close to correct.  
	
<br>
<hr>

## Discussion 

Overall, we synthesized information from three unique datasets that had one common binding: the 50 United States. We then picked out the six most important traits across all three of these data frames, and ranked as well as weighted them based on importance. As discussed above, the word importance is inherently subjective. If one were to choose different weights for these six traits, one would receive a completely different rank for the 50 states. Thus, it is here we will check our assumptions. We assigned these weights based off of feedback from Lynn McGuire, a subject matter expert in environmental conservation, and her experience in the field. By incorporating her feedback as well as key takeaways from the literature review, we hoped to mitigate our bias.

In addition to receiving feedback from subject matter experts, we also ran a multivariate regression test to make sure one could predict rank based off of these scores we assigned to these 6 factors by states. The model proved to be statistically significant and assure us that our math checked out. That being said, all models are inherently wrong; they just give us an idea of approximation. To clarify, this multivariate regression analysis gave us assurance that our math was correct, but does not affirm our ranking system itself is inherently correct.

As for the rankings, the overall order of the results did not surprise us. Washington has the best overall score and the state of Wyoming has the worst score. Here is a breakdown for Washington’s ranks on our 6 factors.

```{r eval=TRUE, echo=FALSE, warning=FALSE, evaluted=FALSE, error=FALSE, message=FALSE}


names(wa.ranks) <- c("Waste", "LFG " ,
                   "Emissions", "Renewables",
                   "Combustables", "Withdrawals")
kable(wa.ranks, caption = "Table 3")
```

Washington ranks number two for both combustables and renewable energy sources. We were curious why this might be, and found that it might have something to do with their laws regarding energy. In 2008, Washington began the Energy Independence Act, which requires large electrical utility companies to use renewable energy. This act sets requirements for these companies that have to do with both conservation and the utilization of renewable energy sources. Companies must attempt to conserve energy in an effective manner, creating a 10 year plan and making 2 year benchmarks. They also must hit renewable energy standards that increase over time, totaling to 15% renewables by 2020 ("Energy Independence", paragraph 2). Regulations like these force the increase the amount of renewable energy generated in the state and are directly responsible for the 76% of energy in the state of Washington being renewable. Ambitious atricles like _A 100% wind, water, sunlgish (WWS) all-sector energy plan for Washington State)_ even claim that Washington state could get 100% of its energy from wind, water, and sunlight by 2050 (Jacobson et. all, 2016, pg. 75)

Here is a breakdown for Wyoming’s score, ranking number 50, on our 6 factors.

```{r eval=TRUE, echo=FALSE, warning=FALSE, evaluted=FALSE, error=FALSE, message=FALSE}

names(wy.ranks) <- c("Waste", "LFG " ,
                   "Emissions", "Renewables",
                   "Combustables", "Withdrawals")
kable(wy.ranks, caption = "Table 4")
```

Looking at Wyoming’s scores we can see that they did poorly in a few different areas. First in the LFG collection they ranked last. This is probably because they don’t have any LFG collection systems at any of their landfills. However, in their defense, they only have a total of 4 landfills in the entire state, the smallest number of landfills in all of the US. If we look at the number of landfills that do not have LFG collection systems for each state, Wyoming actually ties for 4th.

```{r eval=TRUE, echo=FALSE, warning=FALSE, evaluted=FALSE, error=FALSE, message=FALSE}

wy.landfill <- mutate(wy.landfill, num.without = as.character(wy.landfill$num.without))
names(wy.landfill) <- c("State", "Number of Landfills Without LFG Collection", "Percent of Landfills Without LFG Collection")
kable(head(wy.landfill, 8L), caption = "Table 5")
```

This goes to show that if we would have chosen some other way to measure LFG Collection, Wyoming’s scores have changed significantly. It is also further proof that our scores are subjective and it shows the importance of disclosure of our methods and calculations that lead us to these scores.

Another reason why Wyoming is ranked the worst is because Wyoming ranked second to last in the total emissions and didn’t do well in the renewables or combustible power sections as well. This is probably in part due to the fact that they are the number one resource of coal in the United States, mining almost 400 million tons of coal per year ("Coal Mining", paragraph 6). Using coal to produce energy emits large amounts of Sulfur Dioxide (SO2) and Nitrogen Oxides (NOx) ("Air Pollution", 2016, paragraph 1).This is an avoidable problem, as it has been shown management of these wastes at a primary and secondary level can control the pollution via the source (Tiwary, 2001, pg. 1)

<br>
<br>
<hr>

## Conclusion

In conclusion, we created a ranking system that evaluated the 50 United States and assigned them each a score depending on how well they did for each of the 6 variables. Our system for assigning scores proved to be statistically significant on a multivaraite regression model, which simply means our system predicts rank in an appropriate way given the different weights we chose to assign to the six variables. We acknowledge choosing weighs can be subjectective and thus tried to reduce this liability via information from our literature review. Overall, we created a system that offers holistic breadth on environmental impact that another research paper has yet to do. We hope this inspires people to be more conscious of their individual impact on all three facets - landfill, power, and water withdrawal - regardless of their state.  

<br>
<hr>

## Reference List

Access Project and Landfill Data. (2017, October 27). Retrieved December 05, 2017, from https://www.epa.gov/lmop/access-project-and-landfill-data

Air Pollution from Coal Power Plants. (n.d.). Retrieved December 05, 2017, from http://www.greenpeace.org/international/en/campaigns/climate-change/coal/Coal-plants-pollute-the-air/

Anonymous. (2012). North American Power Plant Air Emissions. The Air Pollution Consultant, 22(4), 1_1,1_2,1_3,1_4,1_5,IV.

Coal mining in the United States. (2017, November 20). Retrieved December 05, 2017, from https://en.wikipedia.org/wiki/Coal_mining_in_the_United_States

Emissions & Generation Resource Integrated Database (eGRID). (2017, June 01). Retrieved December 05, 2017, from https://www.epa.gov/energy/emissions-generation-resource-integrated-database-egrid

Energy Independence Act (EIA or I-937). (n.d.). Retrieved December 05, 2017, from http://www.commerce.wa.gov/growing-the-economy/energy/energy-independence-act/

Hodson, E.L., Prinn, D., Martin, R.G., & Hodson, R.G. (2010). The municipal solid waste landfill as a source of ozone-depleting substances in the United States and United Kingdom. Atmospheric Chemistry and Physics, 10(4), 1899-1910.

Jacobson, Delucchi, Bazouin, Dvorak, Arghandeh, Bauer, . . . Yeskoo. (2016). A 100% wind, water, sunlight (WWS) all-sector energy plan for Washington State. Renewable Energy, 86, 75-88.

Mission Areas. (n.d.). Retrieved December 05, 2017, from https://www.usgs.gov/science/mission-areas/water?qt-mission_areas_l2_landing_page_ta=0#qt-mission_areas_l2_landing_page_ta

Tiwary, R.K. Water, Air, & Soil Pollution (2001) 132: 185. https://doi.org/10.1023/A:1012083519667

Wang, H., Small, M., & Dzombak, D. (2014). Factors governing change in water withdrawals for U.S. industrial sectors from 1997 to 2002. Environmental Science & Technology, 48(6), 3420-9.
<br>
<hr>

## Appendix

### Column Tables

#### Landfill
| Column Name                                             | Description                                                                                                                                                                              | 
|---------------------------------------------------------|------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------| 
| Landfill Name                                           | Name of the landfill                                                                                                                                                                     | 
| State                                                   | State that the landfill is in                                                                                                                                                            | 
| Physical Address                                        | Address of the landfill                                                                                                                                                                  | 
| City                                                    | City landfill is located in or near                                                                                                                                                      | 
| County                                                  | County landfill is located in or near                                                                                                                                                    | 
| Zip Code                                                | Zip code for the landfill                                                                                                                                                                | 
| Latitude                                                | Latitude coordinate (decimal) for the landfill                                                                                                                                           | 
| Longitude                                               | Longitude coordinate (decimal) for the landfill                                                                                                                                          | 
| Ownership Type                                          | "Indicates if landfill is publicly owned, privately owned, or co-owned by public and private entities"                                                                                   | 
| Landfill Owner Organization(s)                          | Organization that owns the landfill                                                                                                                                                      | 
| Year Landfill Opened                                    | Year landfill opened or began accepting waste (YYYY)                                                                                                                                     | 
| Landfill Closure Year                                   | Year landfill stopped accepting waste or is expected to stop accepting waste ÿor year landfill closed or is expected to close (YYYY)                                                     | 
| Current Landfill Status                                 | Open/Closed status of landfill                                                                                                                                                           | 
| Waste in Place (tons)                                   | Waste-in-place at the landfill in short tons                                                                                                                                             | 
| Waste in Place Year                                     | Year corresponding to the waste-in-place at the landfill (YYYY)                                                                                                                          | 
| LFG Collection System In Place?                         | "Is there a landfill gas collection system in place? 'Yes' for an active GCCS; 'No' if landfill has no gas collection system, is passively venting/flaring, or has perimeter gas wells." | 
| LFG Collected (mmscfd)                                  | Amount of landfill gas being collected in million standard cubic feet per day                                                                                                            | 
| LFG Flared (mmscfd)                                     | "Amount of landfill gas flared (if project is operational, amount of landfill gas flared in back-up flare(s)) in million standard cubic feet per day"                                    | 
| Current Project Status                                  | "Current project status (Operational, Construction, Planned, Shutdown, Candidate, Potential, Other)"                                                                                     | 
| Project Start Date                                      | Date project became operational (MM/DD/YYYY)                                                                                                                                             | 
| Project Shutdown Date                                   | Date project shut down (MM/DD/YYYY)                                                                                                                                                      | 
| LFG Energy Project Type                                 | Specific type of LFG energy project technology                                                                                                                                           | 
| Project Type Category                                   | Specific type of LFG energy project technology                                                                                                                                           | 
| MW Capacity                                             | Capacity in megawatts for electricity-generating projects                                                                                                                                | 
| Current Year Emission Reductions (MMTCO2e/yr) - Direct  | Direct methane reductions by the energy project for the current year                                                                                                                     | 
| Current Year Emission Reductions (MMTCO2e/yr) - Avoided | Avoided carbon dioxide emission reductions by the energy project for the current year                                                                                                    | 


#### Power
| Column Name                                                      | Column Name ID | Description                                                                                                                                                                                                                                                                                                                                                                            | 
|------------------------------------------------------------------|----------------|----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------| 
| State abbreviation                                               | PSTATABB       | The state_s name abbreviation                                                                                                                                                                                                                                                                                                                                                          | 
| State nameplate capacity (MW)                                    | NAMEPCAP       | It is the intended full-load sustained output of a facility                                                                                                                                                                                                                                                                                                                            | 
| State total annual heat input (MMBtu)                            | STHTIANT       | "The total annual heat input from combustion and noncombustion units, in MMBtu, for the plant. For CHP plants, the value is adjusted by the electric allocation factor."                                                                                                                                                                                                               | 
| State annual net generation (MWh)                                | STNGENAN       | The state_s total reported net generation in MWh.                                                                                                                                                                                                                                                                                                                                      | 
| State ozone season net generation (MWh)                          | STNGENOZ       | "The state_s total five-month ozone season (May through September) net generation in MWh, based on monthly generator generation data. "                                                                                                                                                                                                                                                | 
| State annual NOx emissions (tons)                                | STNOXAN        | "The total annual NOx emissions, in short tons, for the state. Biogas components are adjusted. For CHP plants, the value is adjusted by the electric allocation factor. This adjusted emissions field is estimated by first making the biogas adjustment (if it exists) and then applying the electric allocation factor (if the plant is a CHP). "                                    | 
| State ozone season NOx emissions (tons)                          | STNOXOZ        | "The five-month ozone season (May through September) NOx emissions, in short tons, for the state. Biogas components are adjusted. For CHP plants, the value is adjusted by the electric allocation factor. This adjusted emissions field is estimated by first making the biogas adjustment (if it exists) and then applying the electric allocation factor (if the plant is a CHP). " | 
| State annual SO2 emissions (tons)                                | STSO2AN        | "The total annual SO2 emissions, in short tons, for the state. Landfill gas components are adjusted. For CHP plants, the value is adjusted by the electric allocation factor. This adjusted emissions field is estimated by first making the landfill gas adjustment (if it exists) and then applying the electric allocation factor (if the plant is a CHP). "                        | 
| State annual CO2 emissions (tons)                                | STCO2AN        | "The total annual CO2 emissions, in short tons, for the state. All CO2 emissions from biomass fuels are adjusted to zero. For CHP plants, the value is adjusted by the electric allocation factor. This adjusted emissions field is estimated by first making the biomass adjustment (if it exists) and then applying the electric allocation factor (if the plant is a CHP). "        | 
| State annual CH4 emissions (lbs)                                 | STCH4AN        | "The total annual CH4 emissions, in pounds, for the state. Biogas biomass components are adjusted. For CHP plants, the value is adjusted by the electric allocation factor. This adjusted emissions field is estimated by first making the biomass adjustment (if it exists) and then applying the electric allocation factor (if the plant is a CHP)."                                | 
| State annual N2O emissions (lbs)                                 | STN2OAN        | "The total annual N2O emissions, in pounds for the state. Biogas biomass components are adjusted. For CHP plants, the value is adjusted by the electric allocation factor. This adjusted emissions field is estimated by first making the biomass adjustment (if it exists) and then applying the electric allocation factor (if the plant is a CHP). "                                | 
| State annual coal net generation (MWh)                           | STGENACL       | "The plant annual net generation, in MWh, for coal. Fuel codes that are included in coal are BIT, COG, SUB, LIG, WC, and SC."                                                                                                                                                                                                                                                          | 
| State annual oil net generation (MWh)                            | STGENAOL       | "The plant annual net generation, in MWh, for oil. Fuel codes included in oil are DFO, JF, KER, OO, PC, RFO, RG, and WO."                                                                                                                                                                                                                                                              | 
| State annual gas net generation (MWh)                            | STGENAGS       | "The plant annual net generation, in MWh, for natural gas. Fuel codes included in gas are NG and PG"                                                                                                                                                                                                                                                                                   | 
| State annual nuclear net generation (MWh)                        | STGENANC       | "The plant annual net generation, in MWh, for nuclear. Fuel codes include NUC."                                                                                                                                                                                                                                                                                                        | 
| State annual hydro net generation (MWh)                          | STGENAHY       | "The plant annual net generation, in MWh, for hydro. Fuel codes include WAT."                                                                                                                                                                                                                                                                                                          | 
| State annual biomass net generation (MWh)                        | STGENABM       | "The annual net generation, in MWh, for biomass. Biomass is a fuel derived from organic matter such as wood and paper products, agricultural waste, or methane (e.g., from landfills). The renewable portion of solid waste, fuel code MSB, is included as biomass, as are AB, BLQ, DG, LFG, ME, OBL, OBS, PP, SLW, WDL, and WDS. "                                                    | 
| State annual wind net generation (MWh)                           | STGENAWI       | "The plant annual net generation, in MWh, for wind. Fuel codes include WND."                                                                                                                                                                                                                                                                                                           | 
| State annual solar net generation (MWh)                          | STGENASO       | "The plant annual net generation, in MWh, for solar. Fuel codes include SUN. "                                                                                                                                                                                                                                                                                                         | 
| State annual geothermal net generation (MWh)                     | STGENAGT       | "The plant annual net generation, in MWh, for geothermal. Fuel codes include GEO."                                                                                                                                                                                                                                                                                                     | 
| State annual other fossil net generation (MWh)                   | STGENAOF       | "The plant annual net generation, in MWh, for other fossil fuel that cannot be categorized as coal, oil, or gas. Other fossil fuel codes include BFG, COG, HY, LB, MH, MSF, OG, PRG, and TDF. "                                                                                                                                                                                        | 
| State annual other unknown/ purchased fuel net generation (MWh)  | STGENAOP       | "The plant annual net generation, in MWh, for other unknown/purchased. Fuel codes include OTH, PUR, or WH. "                                                                                                                                                                                                                                                                           | 
| State annual total nonrenewables net generation (MWh)            | STGENATN       | "The annual total nonrenewables net generation, in MWh, for the plant. Nonrenewables are exhaustible energy resources such as coal, oil, gas, other fossil, nuclear power, and other unknown/purchased fuel. This field is the sum of STGENACL, STGENAOL, STGENAGS, STGENAOF, STGENANC, and STGENAOP."                                                                                 | 
| State annual total renewables net generation (MWh)               | STGENATR       | "The annual total renewables net generation, in MWh, for the plant. Renewables are inexhaustible energy resources such as biomass, wind, solar, geothermal, and hydro. This field is the sum of STGENABM, STGENAWI, STGENASO, STGENAGT, and STGENAHY."                                                                                                                                 | 
| State annual total nonhydro renewables net generation (MWh)      | STGENATH       | "The annual total nonhydro renewables net generation, in MWh, for the plant. This field is the sum of STGENABM, STGENAWI, STGENASO, and STGENAGT."                                                                                                                                                                                                                                     | 
| State annual total combustion net generation (MWh)               | STGENACY       | "The annual total combustion net generation, in MWh, for the plant. This field is the sum of STGENACL, STGENAOL, STGENAGS, STGENAOF, STGENABM, and STGENAOP."                                                                                                                                                                                                                          | 
| State annual total noncombustion net generation (MWh)            | STGENACN       | "The annual total noncombustion net generation, in MWh, for the plant. This field is the sum of STGENANC, STGENAHY, STGENAWI, STGENASO, and STGENAGT."                                                                                                                                                                                                                                 | 

#### Water

| Column Name                    | Description                                                          |  | 
|--------------------------------|----------------------------------------------------------------------|--| 
| State                          | The state_s name                                                     |  | 
| Population2010                 | The population in 2010                                               |  | 
| Total-Fresh                    | "Total withdrawals, fresh, in Mgal/d"                                |  | 
| Total-Saline                   | "Total withdrawals, saline, in Mgal/d"                               |  | 
| Total                          | "Total withdrawals, total (fresh+saline), in Mgal/d"                 |  | 
| Groundwater-Fresh              | "Total groundwater withdrawals, fresh, in Mgal/d"                    |  | 
| Groundwater-Saline             | "Total groundwater withdrawals, saline, in Mgal/d"                   |  | 
| Groundwater-Total              | "Total groundwater withdrawals, total (fresh+saline), in Mgal/d"     |  | 
| Surfacewater-Fresh             | "Total surface-water withdrawals, fresh, in Mgal/d"                  |  | 
| Surfacewater-Saline            | "Total surface-water withdrawals, saline, in Mgal/d"                 |  | 
| Surfacewater-Total             | "Total surface-water withdrawals, total (fresh+saline), in Mgal/d"   |  | 
| PublicSuply                    | "Public Supply, total population served, in thousands"               |  | 
| Self-SuppliedDomestic          | "Domestic, self-supplied population, in thousands"                   |  | 
| Irrigation                     | "Irrigation, total withdrawals, fresh, in Mgal/d"                    |  | 
| LiveStock                      | "Livestock, total withdrawals, fresh, in Mgal/d"                     |  | 
| Aquaculture                    | "Aquaculture, total withdrawals, total (fresh+saline), in Mgal/d"    |  | 
| SelfSuppliedIndustrial-Saline  | "Industrial, self-supplied total withdrawals, saline, in Mgal/d"     |  | 
| SelfSuppliedIndustrial-Fresh   | "Industrial, self-supplied total withdrawals, fresh, in Mgal/d"      |  | 
| Mining-Fresh                   | "Mining, total withdrawals, fresh, in Mgal/d"                        |  | 
| Mining-Saline                  | "Mining, total withdrawals, saline, in Mgal/d"                       |  | 
| ThermoelectricPower-Saline     | "Thermoelectric, total withdrawals, saline, in Mgal/d"               |  | 
| ThermoelectricPower-Fresh      | "Thermoelectric, total withdrawals, fresh, in Mgal/d"                |  | 
| Population-Total               | "Total population of county, in thousands"                           |  | 
| Population-Served              | "Public Supply, total population served, in thousands"               |  | 
| Population-Served-Percent      | Population served over the total population                          |  | 
| PublicWithdrawals-Groundwater  | "Public Supply, groundwater withdrawals, total, in Mgal/d"           |  | 
| PublicWithdrawals-Surfacewater | "Public Supply, surface-water withdrawals, total, in Mgal/d"         |  | 
| PublicWithdrawls-Total         | "Public Supply, total withdrawals, total (fresh+saline), in Mgal/d"  |  | 
| WaterDomesticUse               | "Domestic, total use (withdrawals + deliveries)"                     |  | 
| DomesticUse-Percent            | Domestic use over the total population                               |  | 
| AllOtherUse                    | Other uses other than domestic or public supplied                    |  | 
| Irrigation-Groundwater         | "Irrigation, groundwater withdrawals, fresh, in Mgal/d"              |  | 
| Irigation-Surfacewater         | "Irrigation, surface-water withdrawals, fresh, in Mgal/d"            |  | 
| Irigation-Total                | "Irrigation, total withdrawals, fresh, in Mgal/d"                    |  | 
| Livestock-Groundwater          | "Livestock, groundwater withdrawals, fresh, in Mgal/d"               |  | 
| Livestock-Surfacewater         | "Livestock, surface-water withdrawals, fresh, in Mgal/d"             |  | 
| Livestock-Total                | "Livestock, total withdrawals, fresh, in Mgal/d"                     |  | 
| Aquaculture-Groundwater        | "Aquaculture, groundwater withdrawals, total, in Mgal/d"             |  | 
| Aquaculture-Surfacewater       | "Aquaculture, surface-water withdrawals, total, in Mgal/d"           |  | 
| Aquaculture-Total              | "Aquaculture, total withdrawals, total (fresh+saline), in Mgal/d"    |  | 
| Mining-Groundwater             | "Mining, groundwater withdrawals, total, in Mgal/d"                  |  | 
| Mining-Surfacewater            | "Mining, surface-water withdrawals, total, in Mgal/d"                |  | 
| Mining-Total                   | "Mining, total withdrawals, total (fresh+saline), in Mgal/d"         |  | 
| Thermoelectric-Groundwater     | "Thermoelectric, groundwater withdrawals, total, in Mgal/d"          |  | 
| Thermoelectric-Surfacewater    | "Thermoelectric, surface-water withdrawals, total, in Mgal/d"        |  | 
| Thermoelectric-Total           | "Thermoelectric, total withdrawals, total (fresh+saline), in Mgal/d" |  | 
| ThermoelectricPower-Generated  | "Thermoelectric, power generated, in gigawatt-hours"                 |  | 

